 extracting environmenal and oceanographic variables

### 2023-09-13 03_environ_var.Rmd
Update 03_environ_var.Rmd
created env# variable for each buoy, reran with better GPS pts

@@

Sea Surface Temperature from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#tracks_SS_07 <- matchEnvData(tracks_SS_07, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST07", var = c('analysed_sst'))
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST08", var = c('analysed_sst'))
#tracks_SS_10 <- matchEnvData(tracks_SS_10, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST10", var = c('analysed_sst'))
#tracks_SS_12 <- matchEnvData(tracks_SS_12, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST12", var = c('analysed_sst'))
#tracks_SS_13 <- matchEnvData(tracks_SS_13, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST13", var = c('analysed_sst'))
#tracks_SS_14 <- matchEnvData(tracks_SS_14, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST14", var = c('analysed_sst'))
#tracks_SS_16 <- matchEnvData(tracks_SS_16, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST16", var = c('analysed_sst'))
#tracks_SS_18 <- matchEnvData(tracks_SS_18, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST18", var = c('analysed_sst'))
#tracks_SS_19 <- matchEnvData(tracks_SS_19, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST19", var = c('analysed_sst'))
#tracks_SS_20 <- matchEnvData(tracks_SS_20, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST20", var = c('analysed_sst'))
#tracks_SS_21 <- matchEnvData(tracks_SS_21, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST21", var = c('analysed_sst'))
#tracks_SS_22 <- matchEnvData(tracks_SS_22, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST22", var = c('analysed_sst'))
#tracks_SS_23 <- matchEnvData(tracks_SS_23, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST23", var = c('analysed_sst'))
```

@@

for (ss in SSTlist) {
  data <- matchEnvData(get(paste0("tracks_SS_", substr(ss, 33, 34))), nc = ss) %>%
  rename(., sst_mean = analysed_sst_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)   # rename new column to "sst_mean" and remove unnecessary columns

  assign(paste0("tracks_SS_", substr(ss, 33, 34)), data)
}
```
### Wind Stress Curl from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#tracks_SS_07 <- matchEnvData(tracks_SS_07, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress07.nc", var = c('curl'))
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress08.nc", var = c('curl'))
#tracks_SS_10 <- matchEnvData(tracks_SS_10, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress10.nc", var = c('curl'))
#tracks_SS_12 <- matchEnvData(tracks_SS_12, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress12.nc", var = c('curl'))
#tracks_SS_13 <- matchEnvData(tracks_SS_13, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress13.nc", var = c('curl'))
#tracks_SS_14 <- matchEnvData(tracks_SS_14, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress14.nc", var = c('curl'))
#tracks_SS_16 <- matchEnvData(tracks_SS_16, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress16.nc", var = c('curl'))
#tracks_SS_18 <- matchEnvData(tracks_SS_18, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress18.nc", var = c('curl'))
#tracks_SS_19 <- matchEnvData(tracks_SS_19, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress19.nc", var = c('curl'))
#tracks_SS_20 <- matchEnvData(tracks_SS_20, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress20.nc", var = c('curl'))
#tracks_SS_21 <- matchEnvData(tracks_SS_21, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress21.nc", var = c('curl'))
#tracks_SS_22 <- matchEnvData(tracks_SS_22, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress22.nc", var = c('curl'))
#tracks_SS_23 <- matchEnvData(tracks_SS_23, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress23.nc", var = c('curl'))

# Pulls from downloaded file
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'data-raw/erdQMstress1day_LonPM180.nc')
#tracks_SS_08

#windcurl <- erddapToEdinfo('erdQMstress1day', baseurl = 'https://coastwatch.pfeg.noaa.gov/erddap/')

#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc=windcurl, fileName='windcurl.nc')

```
Once downloaded, matches wind stress curl from file
```{r}
windlist <- dir("D:/Environ_variables", recursive=TRUE, full.names=TRUE, pattern="windstress")

for (w in windlist) {
  data <- matchEnvData(get(paste0("tracks_SS_", substr(w, 32, 33))), nc = w) %>%
     dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean, -matchDepth_mean)   # remove unnecessary columns

  assign(paste0("tracks_SS_", substr(w, 32, 33)), data)
}
```

### Distance to continental slope (uses terra package)
```{r}
bathy <- rast("D:/CCES_gebco_bathy.tif") # Bathymetry for all of CCE from https://download.gebco.net/
agg <- terra::aggregate(bathy, 5)  # Increase cell size by factor of 5 to be visible

@@

tracklist <- grep(pattern = "tracks_SS_", names(.GlobalEnv), value = T)

@@

  assign(paste0("tracks_SS_", substr(t, 11, 12)), dbind)
}
```

Depth from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#tracks_SS_07 <- matchEnvData(tracks_SS_07, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth07.nc", var = c('z'))
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth08.nc", var = c('z'))
#tracks_SS_10 <- matchEnvData(tracks_SS_10, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth10.nc", var = c('z'))
#tracks_SS_12 <- matchEnvData(tracks_SS_12, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth12.nc", var = c('z'))
#tracks_SS_13 <- matchEnvData(tracks_SS_13, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth13.nc", var = c('z'))
#tracks_SS_14 <- matchEnvData(tracks_SS_14, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth14.nc", var = c('z'))
#tracks_SS_16 <- matchEnvData(tracks_SS_16, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth16.nc", var = c('z'))
#tracks_SS_18 <- matchEnvData(tracks_SS_18, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth18.nc", var = c('z'))
#tracks_SS_19 <- matchEnvData(tracks_SS_19, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth19.nc", var = c('z'))
#tracks_SS_20 <- matchEnvData(tracks_SS_20, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth20.nc", var = c('z'))
#tracks_SS_21 <- matchEnvData(tracks_SS_21, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth21.nc", var = c('z'))
#tracks_SS_22 <- matchEnvData(tracks_SS_22, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth22.nc", var = c('z'))
#tracks_SS_23 <- matchEnvData(tracks_SS_23, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth23.nc", var = c('z'))
```

@@

for (dd in depthlist) {
  data <- matchEnvData(get(paste0("tracks_SS_", substr(dd, 27, 28))), nc = dd) %>%
  rename(., depth = z_mean) %>%        # rename new column to "depth"
  mutate(depth = depth*-1) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)

  assign(paste0("tracks_SS_", substr(dd, 27, 28)), data)
}
```

#### Bathymetric slope (skipping loop for now)
```{r}
bathy_slope <- terra::terrain(bathy, v = "slope")

for (t in tracklist) {
  slope <- terra::extract(bathy_slope, get(t)[,4:3]) %>%  # Reads latitude and longitude of buoy dataframe
  dplyr::select(-ID) %>%
  rename("bathy_slope" = slope)

@@

  assign(paste0("tracks_SS_", substr(t, 11, 12)), sbind)
}
```

Distance to Shore from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#tracks_SS_07 <- matchEnvData(tracks_SS_07, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore07.nc", var = c('dist'))
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore08.nc", var = c('dist'))
#tracks_SS_10 <- matchEnvData(tracks_SS_10, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore10.nc", var = c('dist'))
#tracks_SS_12 <- matchEnvData(tracks_SS_12, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore12.nc", var = c('dist'))
#tracks_SS_13 <- matchEnvData(tracks_SS_13, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore13.nc", var = c('dist'))
#tracks_SS_14 <- matchEnvData(tracks_SS_14, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore14.nc", var = c('dist'))
#tracks_SS_16 <- matchEnvData(tracks_SS_16, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore16.nc", var = c('dist'))
#tracks_SS_18 <- matchEnvData(tracks_SS_18, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore18.nc", var = c('dist'))
#tracks_SS_19 <- matchEnvData(tracks_SS_19, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore19.nc", var = c('dist'))
#tracks_SS_20 <- matchEnvData(tracks_SS_20, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore20.nc", var = c('dist'))
#tracks_SS_21 <- matchEnvData(tracks_SS_21, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore21.nc", var = c('dist'))
#tracks_SS_22 <- matchEnvData(tracks_SS_22, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore22.nc", var = c('dist'))
#tracks_SS_23 <- matchEnvData(tracks_SS_23, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore23.nc", var = c('dist'))
```

@@

for (d2 in distlist) {
  data <- matchEnvData(get(paste0("tracks_SS_", substr(d2, 32, 33))), nc = d2) %>%
  rename(., dist2shore = dist_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)

  assign(paste0("tracks_SS_", substr(d2, 32, 33)), data)
}
```

### Chlorophyll from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#tracks_SS_07 <- matchEnvData(tracks_SS_07, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll07.nc", var = c('chlorophyll'))
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll08.nc", var = c('chlorophyll'))
#tracks_SS_10 <- matchEnvData(tracks_SS_10, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll10.nc", var = c('chlorophyll'))
#tracks_SS_12 <- matchEnvData(tracks_SS_12, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll12.nc", var = c('chlorophyll'))
#tracks_SS_13 <- matchEnvData(tracks_SS_13, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll13.nc", var = c('chlorophyll'))
#tracks_SS_14 <- matchEnvData(tracks_SS_14, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll14.nc", var = c('chlorophyll'))
#tracks_SS_16 <- matchEnvData(tracks_SS_16, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll16.nc", var = c('chlorophyll'))
#tracks_SS_18 <- matchEnvData(tracks_SS_18, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll18.nc", var = c('chlorophyll'))
#tracks_SS_19 <- matchEnvData(tracks_SS_19, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll19.nc", var = c('chlorophyll'))
#tracks_SS_20 <- matchEnvData(tracks_SS_20, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll20.nc", var = c('chlorophyll'))
#tracks_SS_21 <- matchEnvData(tracks_SS_21, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll21.nc", var = c('chlorophyll'))
#tracks_SS_22 <- matchEnvData(tracks_SS_22, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll22.nc", var = c('chlorophyll'))
#tracks_SS_23 <- matchEnvData(tracks_SS_23, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll23.nc", var = c('chlorophyll'))
```

@@

for (ch in chlalist) {
  data <- matchEnvData(get(paste0("tracks_SS_", substr(ch, 33, 34))), nc = ch) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)

  assign(paste0("tracks_SS_", substr(ch, 33, 34)), data)
}
```

Sea Surface Height Anomalies from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#tracks_SS_07 <- matchEnvData(tracks_SS_07, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh07.nc", var = c('sla'))
#tracks_SS_08 <- matchEnvData(tracks_SS_08, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh08.nc", var = c('sla'))
#tracks_SS_10 <- matchEnvData(tracks_SS_10, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh10.nc", var = c('sla'))
#tracks_SS_12 <- matchEnvData(tracks_SS_12, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh12.nc", var = c('sla'))
#tracks_SS_13 <- matchEnvData(tracks_SS_13, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh13.nc", var = c('sla'))
#tracks_SS_14 <- matchEnvData(tracks_SS_14, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh14.nc", var = c('sla'))
#tracks_SS_16 <- matchEnvData(tracks_SS_16, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh16.nc", var = c('sla'))
#tracks_SS_18 <- matchEnvData(tracks_SS_18, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh18.nc", var = c('sla'))
#tracks_SS_19 <- matchEnvData(tracks_SS_19, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh19.nc", var = c('sla'))
#tracks_SS_20 <- matchEnvData(tracks_SS_20, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh20.nc", var = c('sla'))
#tracks_SS_21 <- matchEnvData(tracks_SS_21, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh21.nc", var = c('sla'))
#tracks_SS_22 <- matchEnvData(tracks_SS_22, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh22.nc", var = c('sla'))
#tracks_SS_23 <- matchEnvData(tracks_SS_23, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh23.nc", var = c('sla'))
```

>>>>

Read in GPSwhale rdas
```{r}
env07 <- readRDS(file = "data/GPSwhaleAIS7.rda")
env08 <- readRDS(file = "data/GPSwhaleAIS8.rda")
env10 <- readRDS(file = "data/GPSwhaleAIS10.rda")
env12 <- readRDS(file = "data/GPSwhaleAIS12.rda")
env13 <- readRDS(file = "data/GPSwhaleAIS13.rda")
env14 <- readRDS(file = "data/GPSwhaleAIS14.rda")
env16 <- readRDS(file = "data/GPSwhaleAIS16.rda")
env18 <- readRDS(file = "data/GPSwhaleAIS18.rda")
env19 <- readRDS(file = "data/GPSwhaleAIS19.rda")
env20 <- readRDS(file = "data/GPSwhaleAIS20.rda")
env21 <- readRDS(file = "data/GPSwhaleAIS21.rda")
env22 <- readRDS(file = "data/GPSwhaleAIS22.rda")
env23 <- readRDS(file = "data/GPSwhaleAIS23.rda")
```

@@

Sea Surface Temperature from ERDDAP using PAMPal function
Daily Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(GPSwhaleAIS7, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST07", var = c('analysed_sst'))
#env08 <- matchEnvData(GPSwhaleAIS8, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST08", var = c('analysed_sst'))
#env10 <- matchEnvData(GPSwhaleAIS10, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST10", var = c('analysed_sst'))
#env12 <- matchEnvData(GPSwhaleAIS12, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST12", var = c('analysed_sst'))
#env13 <- matchEnvData(GPSwhaleAIS13, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST13", var = c('analysed_sst'))
#env14 <- matchEnvData(GPSwhaleAIS14, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST14", var = c('analysed_sst'))
#env16 <- matchEnvData(GPSwhaleAIS16, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST16", var = c('analysed_sst'))
#env18 <- matchEnvData(GPSwhaleAIS18, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST18", var = c('analysed_sst'))
#env19 <- matchEnvData(GPSwhaleAIS19, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST19", var = c('analysed_sst'))
#env20 <- matchEnvData(GPSwhaleAIS20, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST20", var = c('analysed_sst'))
#env21 <- matchEnvData(GPSwhaleAIS21, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST21", var = c('analysed_sst'))
#env22 <- matchEnvData(GPSwhaleAIS22, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST22", var = c('analysed_sst'))
#env23 <- matchEnvData(GPSwhaleAIS23, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST23", var = c('analysed_sst'))
```

@@

for (ss in SSTlist) {
  data <- matchEnvData(get(paste0("env", substr(ss, 33, 34))), nc = ss) %>%
  rename(., sst_mean = analysed_sst_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)   # rename new column to "sst_mean" and remove unnecessary columns

  assign(paste0("env", substr(ss, 33, 34)), data)
}
```
### Wind Stress Curl from ERDDAP using PAMPal function
Wind Stress, Metop-A ASCAT, 0.25°, Global, Near Real Time, 2009-present (1 Day) 
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress07.nc", var = c('curl'))
#env08 <- matchEnvData(env08, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress08.nc", var = c('curl'))
#env10 <- matchEnvData(env10, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress10.nc", var = c('curl'))
#env12 <- matchEnvData(env12, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress12.nc", var = c('curl'))
#env13 <- matchEnvData(env13, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress13.nc", var = c('curl'))
#env14 <- matchEnvData(env14, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress14.nc", var = c('curl'))
#env16 <- matchEnvData(env16, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress16.nc", var = c('curl'))
#env18 <- matchEnvData(env18, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress18.nc", var = c('curl'))
#env19 <- matchEnvData(env19, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress19.nc", var = c('curl'))
#env20 <- matchEnvData(env20, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress20.nc", var = c('curl'))
#env21 <- matchEnvData(env21, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress21.nc", var = c('curl'))
#env22 <- matchEnvData(env22, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress22.nc", var = c('curl'))
#env23 <- matchEnvData(env23, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress23.nc", var = c('curl'))
```
Once downloaded, matches wind stress curl from file
```{r}
windlist <- dir("D:/Environ_variables", recursive=TRUE, full.names=TRUE, pattern="windstress")

for (w in windlist) {
  data <- matchEnvData(get(paste0("env", substr(w, 32, 33))), nc = w) %>%
     dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean, -matchDepth_mean)   # remove unnecessary columns

  assign(paste0("env", substr(w, 32, 33)), data)
}
```

### Distance to continental slope (uses terra and sf package)
```{r}
bathy <- rast("D:/CCES_gebco_bathy.tif") # Bathymetry for all of CCE from https://download.gebco.net/
agg <- terra::aggregate(bathy, 5)  # Increase cell size by factor of 5 to be visible

@@

tracklist <- grep(pattern = "env", names(.GlobalEnv), value = T)

@@

  assign(paste0("env", substr(t, 4, 5)), dbind)
}
```

Depth from ERDDAP using PAMPal function
SRTM15_PLUS Estimated Topography, 15 seconds, Global, v1 
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth07.nc", var = c('z'))
#env08 <- matchEnvData(env08, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth08.nc", var = c('z'))
#env10 <- matchEnvData(env10, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth10.nc", var = c('z'))
#env12 <- matchEnvData(env12, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth12.nc", var = c('z'))
#env13 <- matchEnvData(env13, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth13.nc", var = c('z'))
#env14 <- matchEnvData(env14, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth14.nc", var = c('z'))
#env16 <- matchEnvData(env16, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth16.nc", var = c('z'))
#env18 <- matchEnvData(env18, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth18.nc", var = c('z'))
#env19 <- matchEnvData(env19, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth19.nc", var = c('z'))
#env20 <- matchEnvData(env20, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth20.nc", var = c('z'))
#env21 <- matchEnvData(env21, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth21.nc", var = c('z'))
#env22 <- matchEnvData(env22, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth22.nc", var = c('z'))
#env23 <- matchEnvData(env23, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth23.nc", var = c('z'))
```

@@

for (dd in depthlist) {
  data <- matchEnvData(get(paste0("env", substr(dd, 27, 28))), nc = dd) %>%
  rename(., depth = z_mean) %>%        # rename new column to "depth"
  mutate(depth = depth*-1) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)

  assign(paste0("env", substr(dd, 27, 28)), data)
}
```

#### Bathymetric slope. Units in degrees
```{r}
bathy_slope <- terra::terrain(bathy, v = "slope")

for (t in tracklist) {
  slope <- terra::extract(bathy_slope, get(t)[,3:2]) %>%  # Reads latitude and longitude of buoy dataframe
  dplyr::select(-ID) %>%
  rename("bathy_slope" = slope)

@@

  assign(paste0("env", substr(t, 4, 5)), sbind)
}
```

Distance to Shore from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore07.nc", var = c('dist'))
#env08 <- matchEnvData(env08, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore08.nc", var = c('dist'))
#env10 <- matchEnvData(env10, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore10.nc", var = c('dist'))
#env12 <- matchEnvData(env12, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore12.nc", var = c('dist'))
#env13 <- matchEnvData(env13, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore13.nc", var = c('dist'))
#env14 <- matchEnvData(env14, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore14.nc", var = c('dist'))
#env16 <- matchEnvData(env16, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore16.nc", var = c('dist'))
#env18 <- matchEnvData(env18, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore18.nc", var = c('dist'))
#env19 <- matchEnvData(env19, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore19.nc", var = c('dist'))
#env20 <- matchEnvData(env20, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore20.nc", var = c('dist'))
#env21 <- matchEnvData(env21, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore21.nc", var = c('dist'))
#env22 <- matchEnvData(env22, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore22.nc", var = c('dist'))
#env23 <- matchEnvData(env23, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore23.nc", var = c('dist'))
```

@@

for (d2 in distlist) {
  data <- matchEnvData(get(paste0("env", substr(d2, 32, 33))), nc = d2) %>%
  rename(., dist2shore = dist_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)

  assign(paste0("env", substr(d2, 32, 33)), data)
}
```

### Chlorophyll from ERDDAP using PAMPal function
Chlorophyll-a, Aqua MODIS, NPP, L3SMI, Global, 4km, Science Quality, 2003-
present (Monthly Composite) 
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll07.nc", var = c('chlorophyll'))
#env08 <- matchEnvData(env08, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll08.nc", var = c('chlorophyll'))
#env10 <- matchEnvData(env10, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll10.nc", var = c('chlorophyll'))
#env12 <- matchEnvData(env12, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll12.nc", var = c('chlorophyll'))
#env13 <- matchEnvData(env13, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll13.nc", var = c('chlorophyll'))
#env14 <- matchEnvData(env14, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll14.nc", var = c('chlorophyll'))
#env16 <- matchEnvData(env16, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll16.nc", var = c('chlorophyll'))
#env18 <- matchEnvData(env18, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll18.nc", var = c('chlorophyll'))
#env19 <- matchEnvData(env19, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll19.nc", var = c('chlorophyll'))
#env20 <- matchEnvData(env20, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll20.nc", var = c('chlorophyll'))
#env21 <- matchEnvData(env21, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll21.nc", var = c('chlorophyll'))
#env22 <- matchEnvData(env22, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll22.nc", var = c('chlorophyll'))
#env23 <- matchEnvData(env23, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll23.nc", var = c('chlorophyll'))
```

@@

for (ch in chlalist) {
  data <- matchEnvData(get(paste0("env", substr(ch, 33, 34))), nc = ch) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)

  assign(paste0("env", substr(ch, 33, 34)), data)
}
```

Sea Surface Height Anomalies from ERDDAP using PAMPal function
Sea Surface Height Anomalies from Altimetry, Global, 2017-
present (meters)
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh07.nc", var = c('sla'))
#env08 <- matchEnvData(env08, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh08.nc", var = c('sla'))
#env10 <- matchEnvData(env10, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh10.nc", var = c('sla'))
#env12 <- matchEnvData(env12, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh12.nc", var = c('sla'))
#env13 <- matchEnvData(env13, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh13.nc", var = c('sla'))
#env14 <- matchEnvData(env14, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh14.nc", var = c('sla'))
#env16 <- matchEnvData(env16, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh16.nc", var = c('sla'))
#env18 <- matchEnvData(env18, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh18.nc", var = c('sla'))
#env19 <- matchEnvData(env19, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh19.nc", var = c('sla'))
#env20 <- matchEnvData(env20, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh20.nc", var = c('sla'))
#env21 <- matchEnvData(env21, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh21.nc", var = c('sla'))
#env22 <- matchEnvData(env22, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh22.nc", var = c('sla'))
#env23 <- matchEnvData(env23, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh23.nc", var = c('sla'))
```
